---
title: "bmi_final"
author: "Qianshu Wang"
date: "2017/12/9"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(DT)
library(xtable)
library(knitr)
library(readr)
library(tidyverse)
library(haven)
library(readxl)
library(janitor)
library(pander)
library(Hmisc)
library(multcomp)
library(reshape2)
library(faraway)
library(leaps)
library(HH)
library(broom)
library(GGally)
library(olsrr)
```

### 0. Description
Outcome: length of stay (losdays2 or loshour)

### 1. Clean Data
```{r read_data}
GHP_raw = read_excel("./GHProject_Dataset.xlsx") %>%
  clean_names() 
```

#####1. We may not need to remove ICU_flagged obs.
```{r clean ICU_Flag}
#histogram
GHP_raw %>%
  ggplot(aes(x = losdays2, color = icu_flag)) +
    geom_bar(data = GHP_raw %>% filter(icu_flag == 0), aes(x = losdays2)) +
    geom_bar(data = GHP_raw %>% filter(icu_flag == 1), aes(x = losdays2)) +
    theme_bw()
    
GHP_raw %>%
  ggplot(aes(y = losdays2, x = icu_flag, group = icu_flag)) +
    geom_boxplot()

# 
# GHP = GHP_raw %>%
#   filter(ICU_Flag != 1)
```

#####2. change admitdtm into date formate, relevel cindex and mews based on project instruction, and keep the the first visit for each patient (70 observations were removed)
```{r change date format}
GHP = GHP_raw %>%
  separate(admitdtm, into = c("weekday", "date", "year"), sep = ",") %>% 
  mutate(date = paste(date, " 2016"), date = as.Date(date, format = " %B %d %Y")) %>%
  dplyr::select(-year) %>%
  mutate(cindex = if_else(cindex == 1 & cindex == 2, 1, cindex), 
         cindex = if_else(cindex == 3 & cindex == 4, 2, cindex), 
         cindex = if_else(cindex >= 5, 3, cindex)) %>% # 0 = normal, 1 = mild, 2 = moderate, 3 = severe
  mutate(mews = if_else(mews == 1, 0, mews), 
         mews = if_else((mews == 2) + (mews == 3) == 1, 1, mews), 
         mews = if_else((mews == 4) + (mews == 5) == 1, 2, mews), 
         mews = if_else(mews > 5, 3, mews)) %>% # 0 = normal, 1 = increase caution, 2 = further deterioration, 3 = immediate action required
  group_by(patientid) %>%
  filter(date == min(date)) %>%
  ungroup()
```

#####3. We wanted to investigate the relationship between the length of stay and the health/socioeconomics status of patients, so we decided the variables of intrest to exclude "postalcode", "facilityzip", "facilityname", "date", "weekday", "patientid", and "visitid". Because we also noticed that "loshours" and "losday2" are in fact the same variable with different unit, we chose to only use the "loshours". (why we choose vital signs over mews?)
```{r choose relavent variables}
GHP = GHP %>%
  dplyr::select(-postalcode, -facilityzip, -facilityname, -date, -weekday, -mews, -patientid, -visitid, -losdays2) #%>%
  #mutate(is30dayreadmit = as.factor(is30dayreadmit), icu_flag = as.factor(icu_flag), gender = as.factor(gender), race = as.factor(race), religion = as.factor(religion), maritalstatus = as.factor(maritalstatus), insurancetype = as.factor(insurancetype)) 
  
  # mutate(log_losdays2 = log(losdays2), log_o2sat = log(o2sat), log_resp = log(respirationrate), squre_resp = (respirationrate)^2,  log_bmi = log(bmi), squre_o2sat = (o2sat)^2, log_heart = log(heartrate))
```

#####4. Missing values were replaced with the median of that variable for numeric data. We choose median over mean to address potential skewness in variables.
```{r}
replace_na_w_median = function(vector){
  if (is.numeric(vector)) {
    vector[is.na(vector)] = median(vector, na.rm = TRUE)
  } else if (is.character(vector)) {
    vector[is.na(vector)] = "unknown"
  }
  return(vector)
}

GHP_missing_to_median = map_df(GHP, replace_na_w_median)
```



### 2. Exploratory analysis

#####1. We first examine the distribution of the outcome, "losdays2", with boxplot and histogram.
```{r}
attach(GHP_missing_to_median)
boxplot(loshours)
histogram(loshours)
```
we can see the distribution of the outcome is highly right skewed, so we performed log transformation to it, and the transformed outcome looks symmetrical. 
```{r}
attach(GHP_missing_to_median)
boxplot(log(loshours))
histogram(log(loshours))
GHP_missing_to_median = GHP_missing_to_median %>%
  mutate(log_los = log(loshours)) %>%
  dplyr::select(-loshours)
```
The boxplot suggests 51 outliers in outcome, and they were removed. 
```{r}
attach(GHP_missing_to_median)
lowerbound = quantile(log_los, probs = 0.25) - 1.5 * IQR(log_los)
upperbound = quantile(log_los, probs = 0.75) + 1.5 * IQR(log_los)
GHP_missing_to_median = GHP_missing_to_median %>%
  filter(log_los < upperbound, log_los > lowerbound)
boxplot(log_los)
histogram(log_los)
```

#####2. We use "ggpairs" to investigate the distribution of each variable and check whether they have potential association. (this may goes to the appendix)
```{r ggpairs, message=FALSE, warning=FALSE, fig.height=30, fig.width= 30}
ggpairs(GHP_missing_to_median) + ggplot2::theme_bw()
# attach(GHP)
# boxplot(bpsystolic)
# boxplot(bpdiastolic)
# boxplot(o2sat)
# boxplot(temperature)
# boxplot(heartrate)
# boxplot(respirationrate)
# boxplot(bmi)
# boxplot(ageyear)
# boxplot(cindex)
# # 
# histogram(bpsystolic)
# histogram(bpdiastolic)
# histogram(o2sat)
# histogram(temperature)
# histogram(heartrate)
# histogram(respirationrate)
# histogram(bmi)
# histogram(ageyear)
# histogram(cindex)
# histogram(evisit)
# histogram(icu_flag)
# ggplot(data = GHP_missing_to_median, aes(x = religion)) + geom_bar()
# ggplot(data = GHP_missing_to_median, aes(x = maritalstatus)) + geom_bar()
# ggplot(data = GHP_missing_to_median, aes(x = race)) + geom_bar()
# ggplot(data = GHP_missing_to_median, aes(x = insurancetype)) + geom_bar()
# ggplot(data = GHP_missing_to_median, aes(x = gender)) + geom_bar()
```


### 3. model building

anova tests for categorical data: all categorical variable shows significant difference among their levels, except gender. However, I still include gender in the full model, because...?.
```{r anova for each categorical data}
slr_gender = lm(log_los ~ gender, data = GHP_missing_to_median)
slr_maritalstatus = lm(log_los ~ maritalstatus, data = GHP_missing_to_median)
slr_insurancetype = lm(log_los ~ insurancetype, data = GHP_missing_to_median)
slr_religion = lm(log_los ~ religion, data = GHP_missing_to_median)
anova(slr_gender)
anova(slr_maritalstatus)
anova(slr_insurancetype)
anova(slr_religion)
```

##### 1. The full model is
```{r full model}
# fit all predictor:
mult.reg = lm(log_los ~ bmi + o2sat + heartrate + bpsystolic + bpdiastolic + respirationrate + temperature + is30dayreadmit + cindex + evisit + ageyear + gender + race + religion + maritalstatus + insurancetype, data = GHP_missing_to_median)
  summary(mult.reg)
```
##### 2. Auto
1. stepwise
```{r}
stepwise <- step(mult.reg, direction = 'both')
summary(stepwise)
```
2. forward
```{r}
fwd <- step(mult.reg, direction = 'forward')
summary(fwd)
```
3. backward
```{r}
bwd <- step(mult.reg, direction = 'backward')
summary(bwd)
```
4. reduced model, without marital status
```{r}
auto_reduce1 <- update(stepwise, . ~ . -maritalstatus)
summary(auto_reduce1)
```
5. reduced model, without marital status and o2sat
```{r}
auto_reduce2 <- update(auto_reduce1, . ~ . -o2sat)
summary(auto_reduce2)
```
6. Anova test for nested models
```{r}
anova(auto_reduce1, stepwise)
anova(auto_reduce2, stepwise)
```


##### 3. Criterion
criterion-based variable selection 
```{r}
best <- function(model, ...) 
{
  subsets <- regsubsets(formula(model), model.frame(model), nvmax = 13, ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
}  

best_subsets = best(mult.reg, nbest = 1)
pander(best_subsets)


par(mar=c(4,4,1,1))
par(mfrow=c(1,2))

plot(1:13, best_subsets[, "cp"], xlab="No. of predictor", ylab="Cp Statistic")
abline(0,1)

plot(1:13, best_subsets[, "adjr2"], xlab="No. of predictor", ylab="Adj R2")
```

the selected model
```{r}
crit = lm(log_los ~ heartrate + bpsystolic + bpdiastolic + respirationrate + temperature + is30dayreadmit + cindex + evisit + ageyear + gender + religion + maritalstatus + insurancetype, data = GHP_missing_to_median)
  summary(crit)
```

#check collinearity
```{r}
pander(vif(best_mult))
```

```{r}
best_mult = lm(log_losdays2 ~ log(bmi) + is30dayreadmit + cindex + evisit + ageyear + gender + insurancetype + bpdiastolic + bpsystolic + heartrate +respirationrate+temperature, data = GHP_missing_to_median)
  summary(best_mult)
  
par(mfrow = c(2, 2))
plot(best_mult)

stu_res<-rstandard(best_mult)
outliers_y<-stu_res[abs(stu_res)>2.5]
outliers_y

as.data.frame(influence.measures(best_mult_1))

inf = influence.measures(best_mult)
which(apply(inf$is.inf, 1, any))

ghp_remove <- GHP_missing_to_median[c(-6,-237,-285,-333,-378,-421,-526,-547,-620,-646,-726,-775,-794,-844,-858,-873,-884,-902,-1011,-1147,-1268,-1372,-1427,-1505,-1524,-1551,-1588,-1643,-1678,-1738,-1921,-1929,-2037,-2107,-2190,-2443,-2508,-2576,-2603,-2821,-2837,-2878,-2902,-2935,-2976,-3136,-3154,-3155,-3167,-3182,-3229,-3281,-3357,-3358,-3378,-3389,-3392,-3414,-3462),]

best_mult_1 = lm(log_losdays2 ~ log(bmi) + is30dayreadmit + cindex + evisit + ageyear + gender + insurancetype + bpdiastolic + bpsystolic + heartrate +respirationrate+temperature, data = ghp_remove)
  summary(best_mult_1)
  
inf_1 = influence.measures(best_mult_1)
which(apply(inf_1$is.inf, 1, any))
summary(inf)
  
ghp_outlier = GHP_missing_to_median[c(6,237,285,333,378,421,526,547,620,646,726,775,794,844,858,873,884,902,1011,1147,1268,1372,1427,1505,1524,1551,1588,1643,1678,1738,1921,1929,2037,2107,2190,2443,2508,2576,2603,2821,2837,2878,2902,2935,2976,3136,3154,3155,3167,3182,3229,3281,3357,3358,3378,3389,3392,3414,3462),]

best_mult_2 = lm(log_losdays2 ~ log(bmi) + is30dayreadmit + cindex + evisit + ageyear + gender + insurancetype + bpdiastolic + bpsystolic + heartrate +respirationrate+temperature, data = ghp_outlier)
  summary(best_mult_2)
```


